---
title: "clean_covid-19_confirmed"
author: "Trang Duong"
date: "9/16/2020"
output: html_document
---
---

```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```
```{r general options, include = FALSE}
knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .89, .13, .97)) # plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97)) # plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 1), bty = "n", plt = c(.1, .4, .13, .97))
    else NULL
  },
  margin4 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .18, .95))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })
knitr::opts_chunk$set(margin1 = TRUE, message = FALSE, warning = FALSE,
                      prompt = FALSE, comment = "##", collapse = FALSE,
                      cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
#                      fig.width = 4.24725,
                      fig.width = 4.5,
                      fig.retina = 2, fig.align = "center")
options(width = 263)
l <- "en_US.UTF-8"
Sys.setenv(LANGAGE = l)
Sys.setlocale(locale = l)
Sys.setlocale("LC_MESSAGES", l)
```
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Packages

The packages used in this analysis:

```{r}
library(fitdistrplus)
library(incidence)
library(EpiEstim)
library(mvtnorm)
library(plotrix)
library(car)
# packages from the tidyverse:
library(lubridate)
library(stringr)
library(purrr)
library(tidyr)
library(magrittr) # after tidyr to overload extract()
library(dplyr) # safer to load last
```

## Data sets

The line listing of cases:

```{r}
n_ll <- readRDS("~/Dropbox/Covid-19/covid-19/covid_df.rds") %>% 
  mutate(positive = as_date(ifelse(result1, sample1,
                                   ifelse(result2, sample2,
                                          ifelse(result3, sample3,
                                                 ifelse(result4, sample4,
                                                        ifelse(result5, sample5,
                                                               ifelse(result6, sample6,
                                                                      ifelse(result7, sample7,
                                                                             ifelse(result8, sample8,
                                                                                    ifelse(result9, sample9,
                                                                                           ifelse(result10, sample10,
                                                                                                  ifelse(result11, sample11,
                                                                                                         ifelse(result12, sample12,
                                                                                                                ifelse(result13, sample13,
                                                                                                          
                                                                                                                                    ifelse(result14, sample14,
                                                                                                                              ifelse(result15, sample15,
                                                                                                                                     ifelse(result16, sample16,
                                                                                                                                            ifelse(result17, sample17,
                                                                                                                                                   ifelse(result18, sample18,
                                                                                                                                                          sample19)))))))))))))))))))
                                                                                           , #make sure that the date are in the same format
         entry2onset = as.integer(onset_date - date_last_contact),
         #UPDATE: different formats
         admin2disch = as.integer(date_of_discharge - admission_date),
         admin2onset = as.integer(onset_date - admission_date))
```

## Utilitary functions

### Tuning some functions

Histograms:

```{r}
hist2 <- function(...) hist(main = NA, yaxs = "i", ...)
```

Polygons:

```{r}
polygon2 <- function(...) polygon(border = NA, ...)
```

Plots:

```{r}
plot2 <- function(...) plot(xaxs = "i", yaxs = "i", ...)
```

Arrows:

```{r}
arrows2 <- function(...) arrows(length = .05, angle = 90, code = 3, ...)
```


### Background of figures

The starting dates of each period:

```{r}
periods <- c(
  hubei   = ymd(20200124),
  march15 = ymd(20200315),
  march21 = ymd(20200321),
  april1  = ymd(20200401),
  april23 = ymd(20200423),
  may2    = ymd(20200502)
)
```

The colors for the background:

```{r}
bg <- RColorBrewer::brewer.pal(length(periods) - 1, "YlOrBr")
```

Inverting the last two colors (in order to account for the after-lockdown period):

```{r}
l <- length(bg)
bg <- c(bg[-c(l - 1, l)], bg[l], bg[l - 1])
```

The function that plots the background:

```{r}
plot_layout <- function(ylab = "reproductive number", ylim = c(0, 4), ...) {
  plot2(ymd(20200301), 0, type = "n", xlab = NA, ylab = ylab,
        xlim = c(ymd(20200101) - .5, ymd(20200501) + .5), ylim = ylim,  ...)
  for(i in 1:(length(periods) - 1)) {
    rect(periods[i] - .5, ylim[1], periods[i + 1] - .5, ylim[2], col = bg[i], border = NA)
  }
}
```

### Polygons of cumulated cases

A function that makes the polygons of the cumulative numbers of local transmission:

```{r}
plgn <- function(df, xinf = ymd(20200502), ...) {
  x <- rep(c(df[[1]], xinf), each = 2)
  y <- c(0, rep(df[[2]], each = 2), 0)
  polygon(c(x, x[1]), c(y, y[1]), ...)
}
```


Variables
``` {r}
imported <- covid_df %>%
  filter(infection_type == "imported") %>%
  group_by(quarantine_type == "quarantine center") %>%
  #how do I deal with 44 unknown variables
  tally() %>%
  pull(n)

local <- covid_df %>%
  filter(infection_type != "imported") %>%
  group_by(is.na(infector) == FALSE) %>%
  tally() %>%
  pull(n)
```


The plot:

``` {r}
col <- c("gray93", "grey", "#3182bd", "#9ecae1")
numbers <- c(local, imported)
barplot(matrix(numbers, length(numbers)), col = col, axes = FALSE, xaxs = "i", yaxs = "i")
axis(4)
```

### The timeline of cases
Local transmission:

``` {r}
local <- covid_df %>%
  filter(!is.na(admission_date), infection_type!= "imported") %>%
  group_by(admission_date, infection_type) %>%
  tally() %>%
  ungroup() %>%
  pivot_wider(names_from = infection_type, values_from = n) %>%
  mutate_if(is.integer,replace_na, 0L) %>%
  mutate_if(is.integer, cumsum) %>%
  mutate(admission_date = admission_date - .5 ,#why - .5?,
        `domestic outbreak` = `confirmed case` + inland + contact)
local
```

The plot:

```{r fig.width = 1.75 * 4.24725, margin1 = FALSE, margin2 = TRUE}
plot_layout("cases (n)", c(0, 110))
plgn(local[, c("admission_date", "domestic outbreak")], col = col[1])
plgn(local[, c("admission_date", "contact")], col = col[2])
abline(v = periods[-c(1, length(periods))] - .5, col = bg[-1])
# adding the imported cases:
par(new = TRUE)
covid_df %>% 
  filter(infection_type == "imported") %>% 
  mutate(quarantine_type = ifelse(quarantine_type == "quarantine center", "at home", "treating unit")) %>% 
  group_by(last_contact_date, quarantine_type) %>% 
  tally() %>% 
  pivot_wider(names_from = quarantine_type, values_from = n) %>% 
  left_join(data.frame(last_contact_date = seq(ymd(20200101), ymd(20200501), 1)), ., "last_contact_date") %>% 
  mutate(`no isolation` = ifelse(is.na(`no isolation`) & !is.na(isolation), 0, `no isolation`)) %>% # this last step is needed for the superposition in the barplot that follows
  select(-last_contact_date) %>% 
  t() %>% 
  barplot(col = col[3:4], space = 0, axes = FALSE, ylim = c(0, 110), xaxs = "i", ann = FALSE)
# the legend:
rect(0, 56, 52, 110, col = "white")
#rect(0, 56, 57, 110, col = "white")
l1 <- legend("topleft", legend = c("isolated at entry", "isolated after entry"),
             fill = c(rev(col)), title = expression(bold(" Imported cases:")), title.adj = 0, bty = "n")
l2 <- legend(l1$rect$left, with(l1$rect, top - h),
             legend = c("direct secondary infections", "subsequent secondary infections"),
             fill = c("grey", "gray93"), title = expression(bold(" Cumulative domestic transmissions:")),
             title.adj = 0, bty = "n")
# Grid to help positionning elements:
if(FALSE) {
  abline(v = seq(0, 300, 10))
  abline(h = seq(0, 120, 10))
}
draw.circle2 <- function(...) draw.circle(radius = 1.5, col = "black", ...)
text2 <- function(...) text(adj = .5, cex = .8, col = "white", font = 2, ...)
x <- 30
y <- 8
draw.circle2(x, y)
text2(x, y, "1")
x <- 67.5
y <- 18
draw.circle2(x, y)
text2(x, y, "2")
x <- 77
y <- 33
draw.circle2(x, y)
text2(x, y, "B")
x <- 85
y <- 60
draw.circle2(x, y)
text2(x, y, "H")
x <- 85
y <- 67
draw.circle2(x, y)
text2(x, y, "T")
x <- 97
y <- 100
draw.circle2(x, y)
text2(x, y, "M")
box(bty = "o")
```